---
output: github_document
title:  Example 1 - Loading aimsir17
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

The first step (after installing the package) is to load the library, along with the library dplyr.

```{r,echo=T,message=FALSE}
library(aimsir17)
library(dplyr)
```

There are two datasets that can be used. The first is **observations**

```{r,echo=T}
observations
```

To get see all the variables, use the glimpse() function.

```{r,echo=T}
glimpse(observations)
```

The following variables are recored for each observation:

- **station** (weather station name)
- **year**    (2017)
- **month**   (month of year 1-12)
- **day**     (day of month 1-31)
- **date**    (date object in R)
- **rain**    (hourly rainfall in mm)
- **temp**    (hourly temperature in C)
- **rhum**    (Relative Humidity - percent)
- **msl**     (Mean Sea Level Pressure - hPa)
- **wdsp**    (Mean Wind Speed - knots)
- **wddir**   (Predominant Wind Direction - degree)

The second data set is  **stations**, which can be linked to **observations** via the the variable **station**

```{r,echo=T}
stations %>% print(n=25)
```

The following variables are recored for each station:

- **station**   (weather station name)
- **county**    (location)
- **height**    (height above sea level (m))
- **latitude**  (location coordinates)
- **longitude**  (location coordinates)

A summary of all the observations can be viewed using dplyr.
```{r,echo=T}
observations %>% 
  group_by(station) %>% 
  summarise(TotalRecords=n()) %>%
  print(n=25)
```

A summary of the overall data set can also be viewed. From the summaries, it can be seen that there are s very small number of missing values in the data set.
```{r,echo=T}
summary(observations)
```

