---
output: github_document
title:  Example 3 - Summarising Rainfall Data
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

R tools such as **dplyr** and **tidyr** can be used to summarise data (e.g. add rain observations to obtain monthly and annual cumulative amounts). The three libraries are first loaded.

```{r,echo=T,message=FALSE}
library(aimsir17)
library(dplyr)
library(tidyr)
```

Next, we show the overall data set (219,000 observations)

```{r,echo=T}
observations
```

We can confirm the number of records gathered for each month, and show these
```{r,echo=T}
observations %>% 
  group_by(station, month) %>%
  summarise(TotalObservations=n()) %>%
  pivot_wider(names_from = month,values_from = TotalObservations) %>%
  print(n=25)
```

We can also confirm the number of missing rainfall values for each month.
```{r,echo=T}
observations %>% 
  group_by(station, month) %>%
  summarise(TotalMissing=sum(is.na(rain))) %>%
  pivot_wider(names_from = month,values_from = TotalMissing) %>%
  print(n=25)
```

Finally, we can sum all the rainfall values for each month
```{r,echo=T}
observations %>% 
  group_by(station, month) %>%
  summarise(TotalRainfall=sum(rain,na.rm = T)) %>%
  pivot_wider(names_from = month,values_from = TotalRainfall) %>%
  print(n=25)
```

