---
output: github_document
title:  Example 4 - Exploring Energy Data
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

R tools such as **dplyr** and **tidyr** can be used to summarise data (e.g. add rain observations to obtain monthly and annual cumulative amounts). The three libraries are first loaded.

```{r,echo=T,message=FALSE}
library(aimsir17)
library(dplyr)
library(tidyr)
library(ggplot2)
```

We can show eirgrid's energy data from 2017, which is recorded at 15 minute intervals.

```{r,echo=T}
eirgrid17
```

The variables stored for the eirgrid data include:

```{r,echo=T}
glimpse(eirgrid17)
```
 
The given energy demand for a particular day can be viewed.

```{r,echo=T}
mar17 <- filter(eirgrid17,day==17, month==3)
ggplot(mar17,aes(x=date,y=IEDemand))+geom_point()+geom_line()
```
 
The wind power generated  for a month can be viewed (15 minute intervals)

```{r,echo=T}
mar <- filter(eirgrid17,month==3)
ggplot(mar,aes(x=date,y=IEWindGeneration))+geom_point()+geom_line()
```
 

