---
output: github_document
title:  Example 2 - Exploring Storm Ophelia
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

The data set can be explored for specific events. For example, on October 16th 2017, Storm Ophelia landed in Ireland. We can analyse this data using **dplyr** and **ggplot2**. First, we load the libraries.

```{r,echo=T,message=FALSE}
library(aimsir17)
library(dplyr)
library(ggplot2)
```

Next, we filter the observations for this date.


```{r,echo=T}
o <- observations %>%
  filter(month==10, day==16)
o
```

Next, we can take a selection of stations that had the lowest atmospheric pressure.

```{r,echo=T}
lowest <- o %>% arrange(msl) %>% 
  slice(1:6) %>% 
  pull(station) %>%
  unique()
lowest
```

This is then visualised using ggplot2
```{r,echo=T,fig.width=10, fig.height=6}
ggplot(filter(o,station %in% lowest),aes(x=date,y=msl,colour=station))+
  geom_point()+geom_line()
```


The stations with the highest mean hourly windspeed can be found
```{r,echo=T}
highest<- o %>% arrange(desc(wdsp)) %>% 
  slice(1:6) %>% 
  pull(station) %>%
  unique()
highest
```

```{r,echo=T,fig.width=10, fig.height=6}
ggplot(filter(o,station %in% highest),aes(x=date,y=wdsp,colour=station))+
  geom_point()+geom_line()
```
